On Extensibility and Personalizability of Multi-Modal Trip Planning

### 2016-03-02 paper.tex
Updated sections Implementation and Results, and Conclusion and Future Work.

\usepackage[pdftex]{hyperref}
\usepackage{amsmath}
\usepackage{color}
\usepackage{textcomp}
\usepackage{booktabs}
\usepackage{ccicons}
\usepackage{todonotes}
\usepackage{amsfonts}

@@

\newcommand{\secref}[1]{Section~\ref{sec:#1}}
\newcommand{\figref}[1]{Figure~\ref{fig:#1}}
\newcommand{\tblref}[1]{Table~\ref{tbl:#1}}
\newcommand{\algref}[1]{Algorithm~\ref{alg:#1}}
\newcommand{\thmref}[1]{Theorem~\ref{thm:#1}}
\newcommand{\propref}[1]{Proposition~\ref{prop:#1}}
\newcommand{\corref}[1]{Corollary~\ref{cor:#1}}
\newcommand{\eqtref}[1]{Equation~\ref{eq:#1}}

@@

@@

A natural constraint in trip planning could be ``In this trip I will not drive a car 
after biking or taking the public transit."
In LTL, such constraint can be translated into an LTL formula
\begin{align*}
	\psi = ((M=b) \lor (M=p)) \,\cA\, (\Box (\neg (M=c))).
\end{align*}

@@

We focused on weighted functions over state variables and
designed the cost function, called preferential cost function (PCF), that guides the
graph-based search engine in our trip planner as follows.
\begin{equation}
	\begin{aligned}
		\PCF(S) =& \beta_1 * (\alpha_\walk \cdot T_\walk + \alpha_\bike \cdot T_\bike + \ldots) \\
								&+ (D_\gas + D_\public + \ldots) \\
								&+ \beta_2 * (A_\SUM + \ldots),
	\end{aligned}
	\label{eq:pcf}
\end{equation}
where $\alpha_i$ are real numbers representing the relations among different time pieces,
and $\beta_1$ ($\beta_2$) is the ratio that essentially describes how much in dollars a user would pay to
save an hour (an auxiliary data, respectively).

\subsubsection{Preference Elicitation}
To gather these coefficients ($\alpha_i$'s and $\beta_i$'s) in our $\PCF$, we designed interface to
elicit these numbers from the user.
It asks the user questions and collect answers from the user to derive the coefficients.
These questions are as follows.
\begin{enumerate}
	\item One hr of walking $=$ \underline{\hspace{1cm}} hrs of driving?
	\item One hr of biking $=$ \underline{\hspace{1cm}} hrs of driving?
	\item One hr of public transit $=$ \underline{\hspace{1cm}} hrs of driving?
	\item One hr of taxi $=$ \underline{\hspace{1cm}} hrs of driving?
	\item How much in dollars would you pay to save an hour in traveling?
	\item How much in dollars would you pay to avoid an auxiliary event (e.g., crime) in traveling?
\end{enumerate}
For instance, the user answers 3, 2, 0.25 and 0.5 to the first four questions in the list above.
Intuitively, the numbers indicate that the user prefers public transit the most, followed by
taxi, driving, biking and walking, in order.
We show how we can now derive $\alpha_i$'s in \eqtref{pcf}.
We start with setting $\alpha_\car=1$.
Now, since one hour of walking is equivalent to 3 hours of walking,
we have $\alpha_\walk \times 1 = \alpha_\car \times 3$;
hence, we derive $\alpha_\walk = 3$.
Similarly, we have $\alpha_\bike = 2$, $\alpha_\public = 0.25$,
and $\alpha_\taxi = 0.5$.
Computing $\beta_1$ and $\beta_1$ is straightforward.
Indeed, function $\PCF$ with the input of time, fare and
auxiliary metric pieces boils down to monetary cost,
and the planner computes the best path by optimize
based on this overall monetary cost in the searching
process.

@@

\subsection{System Architecture}
\begin{figure*}[!ht]
  \centering
    \includegraphics[width=0.8\textwidth]{figs/system.pdf}
  \caption{System Overview\label{fig:system}}
\end{figure*}

\subsection{Results}
First, we show resulting route for Alice in \figref{alice}.
\begin{figure}[!ht]
  \centering
    \includegraphics[width=0.5\textwidth]{figs/result_Alice.pdf}
  \caption{Resulting route for Alice\label{fig:alice}}
\end{figure}

Second, we show resulting route for Bob in \figref{bob}.
\begin{figure}[!ht]
  \centering
    \includegraphics[width=0.5\textwidth]{figs/result_Bob.pdf}
  \caption{Resulting route for Bob\label{fig:bob}}
\end{figure}

Third, we show resulting route for Cal in \figref{cal}.
\begin{figure}[!ht]
  \centering
    \includegraphics[width=0.5\textwidth]{figs/result_Cal.pdf}
  \caption{Resulting route for Cal\label{fig:cal}}
\end{figure}

\section{Conclusion and Future Work}

>>>>

\usepackage[pdftex]{hyperref}
\usepackage{amsmath}
\usepackage{color}
\usepackage{textcomp}
\usepackage{booktabs}
\usepackage{ccicons}
\usepackage{todonotes}
\usepackage{amsfonts}
\usepackage{framed}

@@

\newcommand{\secref}[1]{Section~\ref{sec:#1}}
\newcommand{\figref}[1]{Figure~\ref{fig:#1}}
\newcommand{\tblref}[1]{Table~\ref{tbl:#1}}
\newcommand{\algref}[1]{Algorithm~\ref{alg:#1}}
\newcommand{\thmref}[1]{Theorem~\ref{thm:#1}}
\newcommand{\propref}[1]{Proposition~\ref{prop:#1}}
\newcommand{\corref}[1]{Corollary~\ref{cor:#1}}
\newcommand{\eqtref}[1]{Equation~\ref{eqt:#1}}

@@

Our paper is organized as follows.
In the next section, we present what it means for a planner to be extensible
and formally define the method to incorporating new metrics into the planner.
In the following section, we discuss the two aspects of personalization
in trip planning: constraints and preferences, and how they are modeled and
reasoned with in the setting of multi-modal trip planning.
We then move on to describe the system structure of our graph-search based planner,
and show results obtained from our planner in various occassions.
Finally, we conclude by outlining some future research directions.

@@

A natural constraint in trip planning could be ``In this trip I will not drive a car 
after biking or taking the public transit."
In LTL, such constraint can be translated into an LTL formula
\begin{equation}
\label{eqt:ex}
	\psi = ((M=b) \lor (M=p)) \,\cA\, (\Box (\neg (M=c))).
\end{equation}

@@

We focused on weighted functions over state variables and
designed the cost function, called preferential cost function (PCF), that guides the
graph-based search engine in our trip planner as follows.
\begin{equation}
	\begin{aligned}
		\PCF(S) =& \beta_1 * (\alpha_\walk \cdot T_\walk + \alpha_\bike \cdot T_\bike + \ldots) \\
								&+ (D_\gas + D_\public + \ldots) \\
								&+ \beta_2 * (A_\SUM + \ldots),
	\end{aligned}
	\label{eqt:pcf}
\end{equation}
where $\alpha_i$ are real numbers representing the relations among different time pieces,
and $\beta_1$ ($\beta_2$) is the ratio that essentially describes how much in dollars a user would pay to
save an hour (an auxiliary data, respectively).

\subsubsection{Preference Elicitation}
To gather these coefficients ($\alpha_i$'s and $\beta_i$'s) in our $\PCF$, we designed interface to
elicit these numbers from the user.
It asks the user questions and collect answers from the user to derive the coefficients.
These questions are as follows.
\begin{enumerate}
	\item One hr of walking $=$ \underline{\hspace{1cm}} hrs of driving?
	\item One hr of biking $=$ \underline{\hspace{1cm}} hrs of driving?
	\item One hr of public transit $=$ \underline{\hspace{1cm}} hrs of driving?
	\item One hr of taxi $=$ \underline{\hspace{1cm}} hrs of driving?
	\item How much in dollars would you pay to save an hour in traveling?
	\item How much in dollars would you pay to avoid an auxiliary event (e.g., crime) in traveling?
\end{enumerate}
For instance, Alice, an agent, answers 3, 2, 0.25 and 0.5 to the first four questions in the list above.
Intuitively, the numbers indicate that she prefers public transit the most, followed by
taxi, driving, biking and walking, in order.
We show how we can now derive $\alpha_i$'s in \eqtref{pcf}.
We start with setting $\alpha_\car=1$.
Now, since one hour of walking is equivalent to 3 hours of walking,
we have $\alpha_\walk \times 1 = \alpha_\car \times 3$;
hence, we derive $\alpha_\walk = 3$.
Similarly, we have $\alpha_\bike = 2$, $\alpha_\public = 0.25$,
and $\alpha_\taxi = 0.5$.
Computing $\beta_1$ and $\beta_1$ is straightforward.
Indeed, function $\PCF$ with the input of time, fare and
auxiliary metric pieces boils down to monetary cost,
and the planner computes the best path by optimize
based on this overall monetary cost in the searching
process.

@@

\subsection{System Architecture}
\begin{figure*}[!ht]
  \centering
    \includegraphics[width=0.8\textwidth]{figs/system.pdf}
  \caption{System Overview\label{fig:system}}
\end{figure*}
We designed and implemented a multi-modal trip planning
system based on a high-performance graph search engine.
The planner allows user uploads, as well as declarative
constraints and preferences.
We now describe the structure of the planning system.

The trip planner takes two types of data as input:
static data and user-specified request.
The static input includes Map Data and Transit
Data.
Map Data describes the map, a directed graph where 
nodes are street corners, bus stops and train stations.
Transit Data is a set of schedules for the buses and trains
On the other hand, a user provides her request, composed of 
three parts.
First, the user enters \tit{from} and \tit{to} locations 
on the map together with day and time of the start of the trip.  
Second, the user may upload her auxiliary metric dataset, e.g., crime rates.  
Lastly, the user specifies her constraints in LTL and preferences as a $\PCF$.
For example, the constraint 
could be ``never walk through a bad neighborhood."  Given these 
inputs, our planner computes an optimal path satisfying 
all the constraints and optimizing the preferences.
Note that the request from the user is encapsulated into
a JSON object.

For instance, the JSON object for the constraint $\psi$ in \eqtref{ex}
is shown in \figref{exJSON}.
\begin{figure}[!ht]
  \centering
    \includegraphics[width=0.5\textwidth]{figs/exampleJSON.png}
  \caption{JSON object for constraint $\psi$ in \eqtref{ex}\label{fig:exJSON}}
\end{figure}

\subsection{Results}
We present and analyze resulting routes for three agents:
Alice, Bob, and Cal.
This is assuming no auxiliary metric datasets uploaded
so that the agents focus on time and fare.
Fixing their where and when information, we show how
agents' different constraints and preferences affect
their optimal routes computed by our planner.
Their where and when are set so that they all plan to
travel from San Jose International Airport (SJC) in
San Jose, to Pier 39 in San Francisco.
Note that the natural constraint $\psi$ in \eqtref{ex} 
is implicitly imposed on all cases.

Agent Alice has only one constraint that she does not 
have a car; thus, driving to her is never available during this trip.
Per her preferences, Alice provides her thoughts as earlier;
that is, public transportation is preferred to taxi, which is
better than biking, preferred to walking.
What's more, she decides that she would sacrifice thirty dollars
to save one hour during the trip.

The resulting route for Alice is included in \figref{alice}.
The solution takes 2 hours 7 minutes and 18 dollars 94 cents.
It is optimal in the sense the combined value of time
and fare is minimal among all paths from SJC to Pier 39.

\begin{figure}[!ht]
  \centering
    \includegraphics[width=0.5\textwidth]{figs/result_Alice.pdf}
  \caption{Resulting route for Alice\label{fig:alice}}
\end{figure}

Like Alice, Bob is constrained that he will not drive a car in his travel,
and his dollar per hour is thirty.
Moreover, Bob makes his mind to have some workout with his bike and
dictates that he will bike for between one and two hours.
He expresses his preferences: biking and public transit are the most
preferred, next is taxi, and the least preferred is walking.
Similarly, he has done so by answering the aforementioned elicitation questions.

The result for Bob is depicted in \figref{bob}.
It spans 2 hours 57 minutes in time with the fare of 29 dollars 17 cents.
It is so, seemingly worse than what Alice achieved, only because Bob has
the constraint that he will for sure bike for one to two hours, and
the preferences that put biking the most satisfying mode.

\begin{figure}[!ht]
  \centering
    \includegraphics[width=0.5\textwidth]{figs/result_Bob.pdf}
  \caption{Resulting route for Bob\label{fig:bob}}
\end{figure}

Agent Cal will also not use a car in the trip.
He affirms that time is greatly valuable and one hour is measured
to 500 dollars, although his budget is restricted to 50.
Cal's preferences are that the most preferred are public transit
and taxi, and the next preferred are walking and biking that
are equivalent.

Refer to \figref{cal} for the optimal path for Cal.
This route, stretching 1 hour 48 minutes, is the most time-saving one 
compared with the previous two, at the price of 49 dollars 91 cents.
This is due to the constraint that Cal can spend up to 50 dollars,
as well as his preferences and vale of time being high.

\begin{figure}[!ht]
  \centering
    \includegraphics[width=0.5\textwidth]{figs/result_Cal.pdf}
  \caption{Resulting route for Cal\label{fig:cal}}
\end{figure}

\subsubsection{Auxiliary Metric}
When the user of the planner is interested in metrics other than
the ones offered already (i.e., time and fare), she might discover
new metrics (e.g., crime rates and pollution statistics), upload
them into the planner, and retrieve optimal plans taking these
metrics into account.  One scenario of this approach is the following.

Our agent needs to commute back home at night.

\begin{figure}[!ht]
  \centering
    \includegraphics[width=0.5\textwidth]{figs/crime.pdf}
  \caption{Optimal route considering crime rates\label{fig:crime}}
\end{figure}

\section{Conclusion and Future Work}
In multi-modal trip planning, it is vital for the planner to enable
users to integrate new metrics of interest into the existing map,
and to express constraints and preferences over which optimal
routes are computed.
In our work, we designed and built a graph-search based, extensible
and personalizable multi-modal trip planner that utilizes linear
temporal logic and preferential cost
function to model user constraints and preferences.

As a note on the future work, we plan to explore techniques in planning
for computing multiple routes that are diverse with bounded difference of costs.
Continuing this direction, we intend to study the problem of learning the
$\PCF$ coefficients using the observations of the decisions the user made among
the computed paths. (It is reasonable to assume that routes picked by the user
are preferred to others, and these observations could contribute to the learning
problem.)
Also interesting is to introduce traffic information into the planner
to support real-time multi-agent concurrent trip planning.
